---
title: "HW8"
output: html_document
date: "2025-04-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r problem 1}
creatinine <- read.csv("creatinine.csv")
model <- lm(creatclear ~ age, data = creatinine)
summary(model)

predict(model, newdata = data.frame(age = 55))

#Part B
coef(model)["age"]

#Part C
pred_40 <- predict(model, newdata = data.frame(age = 40))
pred_60 <- predict(model, newdata = data.frame(age = 60))

actual_40 <- 135
actual_60 <- 112

resid_40 <- actual_40 - pred_40
resid_60 <- actual_60 - pred_60

resid_40
resid_60
```

### Part A

Using a linear regression of creatinine clearance on age, we predict that a 55-year-old will have a clearance rate of approximately 113.723 mL/min, based on the model

### Part B

The slope of the regression model is −0.620, showing that as age increases, creatinine clearance decreases by about 0.620.

### Part C

Compared to the model’s predictions, the 40-year-old has a creatinine clearance rate 12 mL higher than expected, while the 60-year-old is only about 1.4 mL above expected rate. This shows that the 40-year-old has much healthier kidney function.

## Problem 2

```{r problem 2}
covid <- read.csv("covid.csv")
covid$log_deaths <- log(covid$deaths)
italy_data <- subset(covid, country == "Italy")

italy_model <- lm(log_deaths ~ days_since_first_death, data = italy_data)
italy_growth <- coef(italy_model)["days_since_first_death"]
italy_doubling <- log(2) / italy_growth

italy_growth
italy_doubling

confint(italy_model)["days_since_first_death", ]

log(2) / rev(confint(italy_model)["days_since_first_death", ])



#part 2
spain_data <- subset(covid, country == "Spain")
spain_model <- lm(log_deaths ~ days_since_first_death, data = spain_data)

spain_growth <- coef(spain_model)["days_since_first_death"]
spain_doubling <- log(2) / spain_growth
spain_growth
spain_doubling

confint(spain_model)["days_since_first_death", ]
log(2)/ rev(confint(spain_model)["days_since_first_death", ])


#Part 3
library(ggplot2)
ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1.5) +
  labs(
    title = "Daily COVID-19 Deaths Over Time",
    x = "Days Since First Death",
    y = "Daily Reported Deaths",
    color = "Country"
  ) +
  theme_minimal()
```

### Part 1

-   The estimated growth rate is 0.183, with a 95% confidence interval of (0.161, 0.205).

-   The doubling time is estimated at 3.8 days, with a 95% confidence interval of (3.3, 4.4).

### Part 2

For Spain, the exponential regression model estimates a growth rate of 0.276, with a 95% confidence interval of approximately (0.244, 0.308).

The doubling time is estimated about 2.5 days, with a 95% confidence interval of (2.2, 2.9).

Both intervals for both Spain and Italy are statistically significant because they all do not contain 0 in the interval.

## Problem 3

```{r problem 3}
milk <- read.csv("milk.csv")
milk$log_price <- log(milk$price)
milk$log_sales <- log(milk$sales)

model <- lm(log_sales ~ log_price, data = milk)
summary(model)

confint(model)["log_price", ]
```

-   The estimated price elasticity of demand for milk is −1.619. This indicates that for every 1% increase in price, milk sales decrease by approximately 1.62%.

-   To find the price elasticity, I took the log of price and sales to use the power-law equation that was given. Then I ran a linear regression in R with log(sales) and log(prices) and the slope of that line gave me the elasticity.

-   The 95% confidence interval for the elasticity is (−1.779, −1.458). This means we are 95% confident that the elasticity falls within that range, and since the interval doesn't include 0, meaning it is statistically significant.
